---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tread <a href="https://www.gl-eb.me/tread/"><img src="man/figures/logo.svg" align="right" height="139" alt="tread website" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/gl-eb/tread/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gl-eb/tread/actions/workflows/R-CMD-check.yaml)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.10200272.svg)](https://doi.org/10.5281/zenodo.10200272)
<!-- badges: end -->

tread is an R package to extract data from Excel files generated by Tecan Infinite 200 Pro plate readers.

## Installation

You can install tread from [GitHub](https://github.com/gl-eb/tread) with:

``` r
# install.packages("devtools")
devtools::install_github("gl-eb/tread")
```

### Development Version

You can also install the development version.
While it may contain unreleased features, changes or bugfixes, it is also likely to rapidly evolve, potentially breaking your code.
Do not use the development version if you are not willing to deal with this.

``` r
devtools::install_github("gl-eb/tread", ref = "develop")
```

## Hot To Use

Examples of analysis workflows are documented in `vignette("tread")`

```{r loading}
library(tread)
```

```{r example-custom, eval = FALSE}
# parse your own set of measurements
dat <- tparse("path/to/file.xlsx")

# specify sheet of the excel file (default: 1)
dat <- tparse("path/to/file.xlsx", sheet = 3)

# unite multiple segments of a time series of measurements
dat <- tunite("path/to/file.xlsx")

# specify with which sheets to start and how many to include
dat <- tunite("path/to/file.xlsx", segments = 2, start = 1)
```

## Acknowledgements

Tecan® is a trademark of Tecan Group Ltd., Männedorf, Switzerland.
The authors have no affiliation with them other than being users of their products.

The plate reader icon was designed by [kehan](https://github.com/kehantan) and is licensed under [CC0](https://creativecommons.org/publicdomain/zero/1.0/)
