---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tecanr

<!-- badges: start -->
[![R-CMD-check](https://github.com/gl-eb/tecanr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gl-eb/tecanr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

tecanr is an R package to extract data from .xlsx files generated by Tecan Infinite 200 Pro plate readers.
Currently, only measurements with multiple reads per well are supported due to differences in the layout of the data in the spreadsheets.

tecanr is currently in an unfinished state, so use with caution.

Tecan® is a trademark of Tecan Group Ltd., Männedorf, Switzerland.
I have no affiliation with them other than being a user of their products.

## Installation

You can install tecanr from [GitHub](https://github.com/gl-eb/tecanr) with:

``` r
# install.packages("devtools")
devtools::install_github("gl-eb/tecanr")
```

## Example

```{r loading}
library(tecanr)
```

```{r example-data}
# read supplied example file of single timepoint measurements
dat <- tecan_parse(
  system.file(
    "extdata",
    "tecan_single_time_multi_reads.xlsx",
    package = "tecanr"
  )
)
dat
```

```{r example-custom, eval = FALSE}
# read your own set measurements
dat2 <- tecan_parse("path/to/file.xlsx")

# specify sheet of the excel file (default: 1)
dat3 <- tecan_parse("path/to/file.xlsx", sheet = 3)
```

```{r example-data-multi}
# read supplied example file of multiple timepoint measurements
# parameters work the same as for tecan_read_single()
dat4 <- tecan_parse(
  system.file(
    "extdata",
    "tecan_time_series_multi_reads.xlsx",
    package = "tecanr"
  )
)
dat4
```

```{r example-data-temp}
# read temperature over time from example file of multiple timepoint
# measurements
dat_temperature <- tecan_read_temperature(
  system.file(
    "extdata",
    "tecan_time_series_multi_reads.xlsx",
    package = "tecanr"
  )
)
dat_temperature
```
