---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tecanr

<!-- badges: start -->
[![R-CMD-check](https://github.com/gl-eb/tecanr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gl-eb/tecanr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

tecanr is an R package to extract data from Excel files generated by Tecan Infinite 200 Pro plate readers.
Currently, only measurements with multiple reads per well are supported due to differences in the layout of the data in the spreadsheets.

tecanr is currently in an unfinished state, so use with caution.

Tecan® is a trademark of Tecan Group Ltd., Männedorf, Switzerland.
I have no affiliation with them other than being a user of their products.

## Installation

You can install tecanr from [GitHub](https://github.com/gl-eb/tecanr) with:

``` r
# install.packages("devtools")
devtools::install_github("gl-eb/tecanr")
```

### Development Version

You can also install the development version.
While it may contain unreleased features, changes or bugfixes, it is also likely to rapidly evolve, potentially breaking your code.
Do not use the development version if you are not willing to deal with this.

``` r
devtools::install_github("gl-eb/tecanr", ref = "develop")
```

## Hot To Use

```{r loading}
library(tecanr)
```

```{r example-custom, eval = FALSE}
# parse your own set of measurements
dat <- tecan_parse("path/to/file.xlsx")

# specify sheet of the excel file (default: 1)
dat <- tecan_parse("path/to/file.xlsx", sheet = 3)

# unite multiple segments of a time series of measurements
dat <- tecan_unite("path/to/file.xlsx")

# specify with which sheets to start and how many to include
dat <- tecan_unite("path/to/file.xlsx", segments = 2, start = 1)
```

### Examples

```{r example-data}
# parse supplied example file of time series with multiple measurements per well
# at each timepoint
tecan_parse(
  system.file(
    "extdata",
    "tecan_time_series_multiple_reads.xlsx",
    package = "tecanr"
  )
)
```

```{r example-data-multi}
# read supplied example file of a time series of measurements where the plate
# reader program was stopped and restarted, resulting in the data being split
# into multiple segments / sheets
tecan_unite(
  system.file(
    "extdata",
    "tecan_time_series_segments.xlsx",
    package = "tecanr"
  )
)
```
